---
title: "Project_Milestone_04"
author: "Ngan Nguyen, Nicole Fernandez, Shirley Sui"
format: html
editor: visual
---

## Project Milestone #4: Visualizations

Submit an Rmd or Qmd and publish an html file on RPubs with the following:

## Final datasets for creation of visualization

```{r load libraries, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)

# Load tidyverse for all data cleaning packages
library(tidyverse)
```

```{r load dataset}
#Load data from project Milestone #3 in Milestone_3_Clean_Data
df_ca_vax_clean <- read_csv('./Milestone_3_Clean_Data/df_ca_vax_clean.csv')
df_flu_clean <- read_csv('./Milestone_3_Clean_Data/df_flu_clean.csv')
```

-   Join all datasets together

```{Join all datasets together}

```

-   Calculate any remaining data elements needed for analysis
-   Show code used to create joined dataset, but please do not print full data frame output (showing data structure with str() is okay)

## Visualizations (at least one per group member)

-   One print quality table as requested in scenario
-   One print quality plot or chart as requested in scenario
-   For groups of 3, one additional print quality table or plot of your choice (can support the requested data in the scenario, or answer a different question using the same data sources)

## For each visual, include - Code used to generate visual - Legend (if necessary) - 1-2 sentence interpretation

NOTE: Please make sure the visual can stand-alone, meaning it includes enough information in title, legend, and footnote so that a person who sees only the visualization could understand what is being presented. Please also make sure column names, axis labels, and any other labels are meaningful and not just the name of the variable (ex: "County" rather than "county_name")

## Html file that is professionally prepared for presentation and published to RPubs

-   Only the necessary information is in the output (e.g., suppress entire data frame outputs but showing data structure with str() is okay.
-   Code used to import and clean data (milestones 2 and 3) can be included in the rmd/qmd but please use the option include=FALSE in order to prevent the code and any output from. Code chunk would look like {r chunk_name, include=FALSE}
-   Utilize warning=FALSE and message=FALSE to suppress any warnings/messages that may be displaying from any chunks
-   For this milestone please make sure code for visualizations are included in the final file. - Show your work by "echoing" code used in Rmd/Qmd file to create your tables and graphs. \`\`\`{r, echo = TRUE} in code chunk preferences.
-   Use headers and sub headers to create an organized document \## Running Code
