---
title: "Project_Milestone_06"
author: "Ngan Nguyen, Nicole Fernandez, Shirley Sui"
format: html
editor: visual
---

```{r load & set up, include=FALSE, echo=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)

# Load tidyverse for all data cleaning packages
library(tidyverse)
library(kableExtra)
library(ggplot2)
library(plotly)

#Load data from project Milestone #3 in Milestone_3_Clean_Data
df_vax <- read_csv('./Milestone_3_Clean_Data/df_ca_vax_clean_subset.csv')
df_flu <- read_csv('./Milestone_3_Clean_Data/df_flu_clean_aggregate.csv')

# Join all datasets together
# Flu data only has quarters 2022.4 to 2023.2
# Vax data has quarters 2022.1 to 2023.3
# Before joining, filter vax data to only 2022.4 to 2023.2

df_vax <- df_vax %>%
  filter(quarter >= 2022.4 & quarter <= 2023.2)

# perform inner_join - we don't want any data that doesn't have a match in both datasets

df_joined <- inner_join(df_flu, df_vax, by = join_by(age_category, county, quarter)) %>% select(!c(sum_dx_new, sum_severe_new)) %>% drop_na()
```

## [Problem Statement]{.underline}

Background & research question(s)/null & alternative hypotheses here

## [Methods]{.underline}

-   Description of datasets (data source), years and/or dates of data
-   Description of cleaning, creating new variables (data manipulation)
-   Analytic methods

## [Results]{.underline}


```{r covid vax vs flu rate - Shirley, warning = FALSE, message = FALSE, echo = FALSE}
# COVID vaccination rates vs flu rate (new cases), per quarter, color coded by county

# Re-format facet labels for quarters
qtr.labs <- c("Q4-2022", "Q1-2023", "Q2-2023")
names(qtr.labs) <- c(2022.4, 2023.1, 2023.2)

vax_flu_byqtr <- df_joined %>%
  group_by(county, quarter) %>%
  mutate(mean_vaccination_rate = mean(mean_vaccination_rate), dx_new_rate = mean(dx_new_rate)) %>%
  ggplot(aes(x = mean_vaccination_rate, y = dx_new_rate, color=county)) +
  geom_point() +
  geom_smooth(method = lm, color = "red") +
  theme_minimal() +
  scale_y_continuous(labels= function(x) round(x*100, digits = 2)) + 
  labs(x = "Mean COVID Vaccination Rate 
  (fully vaccinated / population) ", y = "Mean Flu Rate 
  (new diagnoses / 100 susceptible)", title = "Figure 1: COVID vaccination rate does not significantly reflect 
flu vaccination rate", subtitle = "based on quarterly COVID vaccination rate data compared to simulated flu incidence rate data per 
quarter across 57 California counties, from Q4-2022 to Q2-2023") +
  facet_wrap(~quarter, scales = "free", labeller = labeller(quarter = qtr.labs)) + 
  theme(
    legend.position = "none", 
    plot.title.position = "plot",
    plot.title = element_text(face = "bold"), 
    plot.subtitle = element_text(face = "italic"), 
    strip.text.x = element_text(face = "bold"),
    axis.title.y = element_text(margin = margin(r = 5)),
    axis.title.x = element_text(margin = margin(t = 5)))

ggplotly(vax_flu_byqtr) %>%
  layout(
    title = list(text = paste0('Figure 1: COVID vaccination rate does not significantly reflect 
flu vaccination rate.','<br>','<sup>','Across 57 California counties, from Q4-2022 to Q2-2023','</sup>'), yref = 'container', xref = 'paper', yref = 'paper'), 
    annotations = list(text = paste0('<sup>','Based on quarterly COVID vaccination rate data compared to simulated flu incidence rate data per quarter','</sup>'), showarrow = F, xref = 'paper', yref = 'paper', xanchor = 'right', yanchor = 'auto', xshift = 350, yshift = -200),
    margin = list(t = 150, b = 110, l = 100, r = 30)
  )
```

```{r Correlation table indicating relationship between COVID vaccination rates on severe flu rate - Kim, warning = FALSE, message = FALSE, echo = FALSE}

# summary(df_joined)

# Create categories based on COVID vaccination rates and severe flu rates
data <- df_joined %>%
  mutate(
    COVID_Vaccination_Rate = ifelse(mean_vaccination_rate <= median(mean_vaccination_rate), "Low Vaccination Rate (Min = 0.14)", "High Vaccination Rate (Max = 1.14)"),
    Severe_Flu_Rate = ifelse(severe_new_rate <= median(severe_new_rate), "Low Severe Flu Rate (Min = 0.00)", "High Severe Flu Rate (Max = 0.43)")
  )

# Calculate the correlation for each combination
correlation_table <- data %>%
  group_by(COVID_Vaccination_Rate, Severe_Flu_Rate) %>%
  summarise(
    Correlation = round(cor(mean_vaccination_rate, severe_new_rate), 2)
  )

# Rename columns
colnames(correlation_table) <- c("COVID Vaccination Rate", "Mean Severe Flu Rate", "Correlation")

# Print the correlation table with kableExtra
  kable(
    correlation_table,
    format = "html", 
    caption = "Table 1: No clear correlation between COVID vaccination rates and mean severe flu rate, Q4-2022 - Q2-2023") %>% 
  kable_styling(full_width = FALSE) 


```

```{r age group vax vs flu severity -Nicole, warning = FALSE, message = FALSE, echo = FALSE}
# Visualization: One print quality plot or chart as requested in scenario
# Using Both a table & a chart seem appropriate for this specific research question
# Rename joined df
age_group_vax_vs_flu_severity <- df_joined

# Rename df
age_group_vax_vs_flu_severity_table <- age_group_vax_vs_flu_severity

# Create a summary table
summary_table <- age_group_vax_vs_flu_severity_table %>%
  group_by(age_category) %>%
  summarize(mean_vaccination_rate = mean(mean_vaccination_rate),
            mean_flu_severity = mean(severe_new_rate))

# Rename summary table data
colnames(summary_table ) <- c("Age Category", "Mean COVID Vaccination Rate", "Mean Severe Flu Rate")

# Print the table using kableExtra (Quality table #1)
kable(
  summary_table, 
  "html",
  caption = "Table 2: High COVID vaccination rate not significantly associated with low severe flu rate") %>%
  kable_styling(full_width = FALSE)


# Quality plot

# Scatter plot with facets for each age category
vax_flu_byage <- ggplot(data = age_group_vax_vs_flu_severity, aes(x = mean_vaccination_rate, y = severe_new_rate, color = mean_vaccination_rate)) +
  geom_point(size = 3) +
  labs(title = "Figure 2: High COVID vaccination rate not significantly associated with
low severe flu rate",
       subtitle = "Scatter Plots by Age Categories",
       x = "Mean Vaccination Rate",
       y = "Flu Severity Rate",
       color = "Mean Vaccination Rate") +
  geom_smooth(method = lm, color = "yellow") +
  scale_color_gradient(low = "blue", high = "red") +  
  facet_wrap(~age_category, scales = "free") +  
  theme_minimal() + 
  theme(panel.margin = unit(2, "lines"))

# vax_flu_byage
# interactive
ggplotly(vax_flu_byage) %>%
  layout(
    title = list(text = paste0('Figure 2: High COVID vaccination rate not significantly associated with
low severe flu rate','<br>','<sup>','Scatter Plots by Age Categories','</sup>'), yref = 'container', xref = 'paper', yref = 'paper'),
    margin = list(t = 130, b = 0, l = 100, r = 30)
  )

```

## [Discussion]{.underline}

** NOTE: I just pasted all of our interpretations here - it should be modified to flow better and maybe provide more context. Some of this might belong more in the Results section, too**

It's crucial to recognize that this analysis pertains to aggregate-level data, and caution should be exercised in generalizing these findings to individual-level dynamics to avoid the ecological fallacy.

Analysis of the COVID vaccination rate data and simulated flu data indicates that COVID vaccination does not seem to strongly reflect flu vaccination within the 57 California counties from which data was collected/simulated. Additional measures to promote flu vaccination may be needed.

Correlation values per each case provide insights into the relationships between COVID vaccination rates and severe flu cases rates in different scenarios, ranging from weak negative to moderate positive correlations. There is a significant weak negative correlation between the high vaccination rate group and low severe flu rate which might be impacted by confounders.

COVID vaccination versus flu severity by age group interpretation: The data indicates an age-related uptrend in vaccination rates among the older demographics. Intriguingly, despite heightened vaccination coverage in these age groups, there is a simultaneous upswing in average flu severity.
